---
title: "NetRes Package Vignette"
author: "Boris Hayete"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Example 1 - star topology network generation and inference

```{r setup, include = TRUE}
##pdf('Example1.pdf')
set.seed(42)
##con <- file("test.log")
##sink(con, append=TRUE)
##sink(con, append=TRUE, type="message")

library(NetRes)
library(bnlearn)

##demo script
network = SFNetwork$new(numVertices=50, topology='star', addOneEdge = FALSE)
generated = network$generateData(numSamples = 100, latIdx = 1)
bnlearn::graphviz.plot(generated$graph, layout='dot', highlight = list(nodes = grep('U_', colnames(generated$data), value=T), col = "tomato", fill = "orange"))

tabuArgs = list(start = NULL, whitelist = NULL, blacklist = NULL, score = 'ebic-g', debug = FALSE, tabu = 10, max.iter = Inf, maxp = Inf, optimized = TRUE)

netResObj = NetRes$new(generated$data, true.graph = generated$graph, nIter = 10, nBoot=100, algorithm='tabu', 
                       algorithm.args = tabuArgs, mode='foo', latentSpaceMethod='VAE') #'sparse.pca') 
##netResObj$assess() #if the true network has 0 edges among the observable variables, that causes crashes
##after inference
bnlearn::graphviz.plot(netResObj$ensemble[[1]][[1]], layout='dot')
bnlearn::graphviz.plot(netResObj$ensemble[[2]][[1]], layout='dot', highlight = list(nodes = grep('U_', bnlearn::nodes(netResObj$ensemble[[2]][[1]]), value=T), col = "tomato", fill = "orange"))
##plot(netResObj)
##    score(net, data, type = algorithm.args$score, prior = algorithm.args$prior)
##sink()
##sink(type='message')
##dev.off()
##rm(netResObj)

```

Example 2 - twin-star topology network generation and inference

```{r setup, include = TRUE}
##pdf('Example2.pdf')
set.seed(42)

##con <- file("test.log")
##sink(con, append=TRUE)
##sink(con, append=TRUE, type="message")

library(NetRes)
library(bnlearn)

##demo script
network = SFNetwork$new(numVertices=200, topology='twin.star', addOneEdge=FALSE)
generated = network$generateData(numSamples = 300, latIdx = c(1, 2)) #used to be 500
print(network)
plot(network)
bnlearn::graphviz.plot(generated$graph, layout='dot', highlight = list(nodes = grep('U_', colnames(generated$data), value=T), col = "tomato", fill = "orange"))

tabuArgs = list(start = NULL, whitelist = NULL, blacklist = NULL, score = 'ebic-g', debug = FALSE, tabu = 10, max.iter = Inf, maxp = Inf, optimized = TRUE)

netResObj = NetRes$new(generated$data, true.graph = generated$graph, nIter = 10, nBoot=100, algorithm='tabu', 
                       algorithm.args = tabuArgs, mode='foo',
                       latentSpaceMethod='sparse.pca') 
##netResObj$assess() #if the true network has 0 edges among the observable variables, that causes crashes
##after inference
bnlearn::graphviz.plot(netResObj$ensemble[[1]][[1]], layout='dot')
bnlearn::graphviz.plot(netResObj$ensemble[[2]][[1]], layout='dot', highlight = list(nodes = grep('U_', bnlearn::nodes(netResObj$ensemble[[2]][[1]]), value=T), col = "tomato", fill = "orange"))
save.image(file='example2.RData')
plot(netResObj)
##rm(netResObj)
##sink()
##sink(type='message')
##dev.off()

```

Example 3 - an example of scale-free network generation and inference
```{r, include=TRUE}
pdf('Example3.pdf')
con <- file("test.log")
sink(con, append=TRUE)
sink(con, append=TRUE, type="message")

set.seed(42)

print('Example 3')
library(NetRes)
library(bnlearn)

##demo script
network = SFNetwork$new(numVertices=1000, numNeighbors = 4)
generated = network$generateData(numSamples=100, latIdx = c(2, 3), rescaleNoiseBy=NULL)
print(network)
bnlearn::graphviz.plot(generated$graph, layout='dot', highlight = list(nodes = grep('U_', colnames(generated$data), value=T), col = "tomato", fill = "orange"))

##algArgs = list(start = NULL, whitelist = NULL, blacklist = NULL, score = 'bge', prior = 'vsp', debug = FALSE, tabu = 100, max.iter = Inf, maxp = Inf, optimized = TRUE, max.rank=10)
algArgs = list(start = NULL, whitelist = NULL, blacklist = NULL, score = 'ebic-g', debug = FALSE, tabu = 100, max.iter = Inf, maxp = Inf, optimized = TRUE, max.rank=10)

netResObj = NetRes$new(generated$data, true.graph = generated$graph, nIter = 10, nBoot=100, algorithm='tabu', algorithm.args = algArgs, mode='normal', weightedResiduals=FALSE, scale=TRUE, latentSpaceMethod='sparse.pca', optimizeLatentSpace=FALSE) 

netResObj$assess()
nIter = length(netResObj$ensemble)
bnlearn::graphviz.plot(netResObj$ensemble[[1]][[1]], layout='dot')
bnlearn::graphviz.plot(netResObj$ensemble[[2]][[1]], layout='dot', highlight = list(nodes = grep('U_', bnlearn::nodes(netResObj$ensemble[[nIter]][[1]]), value=T), col = "tomato", fill = "orange"))

##ENSURE THAT ALL LATENT VARIABLES ARE SCALED PROPORTIONALLY RATHER THAN ALL TO 1 (IF THAT'S THE RIGHT THING - CHECK!!!!).  THIS WOULD BE AT THE END OF THE GA ITERATION.  NEXT THING, TRY ALLOWING GREATER MARGIN FOR VARIABLES WITH COEFFICIENTS NOT SET TO 1.
##Add F1 measure to assess

##plot(netResObj)
##rm(netResObj)

sink()
sink(type='message')
dev.off()

```

Example 4 - a real example of gene expression and protein levels of neurons in the striatum of mice with Huntington's disease
The latent variable is the length of CAG repeats in Huntingtin gene, which is the only genetic marker of age of onset of HD and accounts
for 50% of variability in the age of onset.
Here we don't know the true graph, but we DO know the latent variable.
Note: the authors gratefully acknowledge "Cure Huntington's Disease Foundation" for their permission to use this snippet of their data.
```{r, include=TRUE}
##pdf('Example4.pdf')
###con <- file("test.log")
##sink(con, append=TRUE)
##sink(con, append=TRUE, type="message")
data(striatum) #load the HD dataset
striatum = scale(striatum) #scale all variables

##create a blacklist making Age and Gender input-only variables (these are experimental or, effectively, interventional variables)
##do not include the latent space in the blacklist
library(dplyr)
exptVars =  bnlearn::tiers2blacklist(list(setdiff(colnames(striatum), c('Age', 'Gender')), c('Age', 'Gender'))) %>% as.data.frame %>% dplyr::filter(to != 'U_CAG')

set.seed(42)

library(NetRes)
library(bnlearn)

algArgs = list(start = NULL, whitelist = NULL, blacklist = exptVars, score = 'ebic-g', debug = FALSE, tabu = 100, max.iter = Inf, maxp = Inf, optimized = TRUE, max.rank=10)

netResObj = NetRes$new(striatum, nIter = 10, nBoot=100, algorithm='tabu', algorithm.args = algArgs, mode='normal', weightedResiduals=FALSE, scale=TRUE, latentSpaceMethod='VAE', optimizeLatentSpace=FALSE) 

onetResObj$assess()
nIter = length(netResObj$ensemble)
bnlearn::graphviz.plot(netResObj$ensemble[[1]][[1]], layout='dot')
bnlearn::graphviz.plot(netResObj$ensemble[[2]][[1]], layout='dot', highlight = list(nodes = grep('U_', bnlearn::nodes(netResObj$ensemble[[4]][[1]]), value=T), col = "tomato", fill = "orange"))

##sink()
##sink(type='message')
##dev.off()

```
